\documentclass[11pt, oneside]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{epstopdf}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage[]{mcode}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\prox}{prox}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\diag}{diag}
\DeclareMathOperator*{\Lap}{Lap}
\DeclareMathOperator*{\E}{E}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{STAT 520 - Project Proposal}
\author{Jason Hartford - 81307143 \\
Dustin Johnson - 11338118}
%\date{}                                           % Activate to display a given date or no date

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Outline}

Markov Chain Monte Carlo (MCMC) has become the standard sampling tools that have enabled Bayesian statistics to evolve and be applied to increasingly complex and higher dimensional models. Probabilistic programming languages such as JAGS, BUGS, and Stan have readily integrated MCMC into a "black-box" declarative structure, making Bayesian model construction more simplistic. However, increasing distribution complexity and dimensionality has placed higher demands for improvements in sampling methods in terms of performance and computational efficiency. 
	
Sequential Monte Carlo (SMC) samplers offer an alternative to MCMC by offering the ablity to sample from a posterior distribution through a novel online algorithm, as opposed to sampling from the entire joint distribution each time new observations enters the model. Therefore, SMC improves computational speed and adaptability of the posterior estimate with new data. The downside of this approach is the possible difficulty of constructing the Markov transition probabilities to conform to the desired distribution.

########## so far ^

MCMC is the standard method of sampling from distributions with high dim stuff...  used in JAGS, BUGS, Stan, etc.

SMC samplers offer the promise of better performance in .... situations, with the added difficulty of having to come up with transitions from simple to desired distribitions.

But... relatively little is know about their relative performance. Given the added cost to practitioners of coming up with transition blah blah blah, the natural question to ask is how much ``performance" can be gained by using SMC samplers.

To test this we will consider the performance of MCMC using Metro Hastings and Gibbs sampling, and SMC samplers in these popular contexts: bayesian non-parametric, directlet process, gaussian process, mixture models, and hierarchical models.

% MCMC - Metro Hasting
% MCMC - Gibbss
% 


\end{document}  